# Composing tweets {#composing}

The most common question I get at social media workshops that I've run is, "What should I tweet?". There is no easy answer to this, because every researcher and their subfield is different. But regardless of your subfield, __the best way to engage your followers is to either entertain or educate__. In other words, you should aim to help people either pass time or save time. Being entertaining doesn't come naturally to most people, so don't worry if this isn't you. But as a scientist, you are very well-placed to educate, no matter your level of training.

Thanks to [Chapter 1](#beginner), you're up to speed with the general mechanics of Twitter. You also just learned about the "entertain or educate" principle, with *any* scientist ready to educate. But even when arriving at this point, many scientists hit a wall, because they believe they can only tweet the finished product of their work, which is typically papers. But as soon as you realise that the *process* of your work is just as interesting as the *output*, then things become much easier. You're going to see a few examples of this below.

When it comes to tweet frequency, it's hard to give any firm recommendations. Whether you're tweeting too much depends on how many people your followers are following, and how much *they* tweet. So, this means that the "too much" threshold is different for everyone. In my opinion, I think it's fairly difficult to tweet too much. In the early days of Twitter (or any social network), where users didn't follow that many people, it was pretty easy to flood someone's timeline. But now, people tend to follow hundreds (sometimes thousands) of accounts, so this is harder to do. I think the upside of more tweets outweighs the downside a few people unfollowing you because they think you tweet too much. You never know which tweets other people will find interesting.

## Some example tweets {.unnumbered}

### Tweeting about your own research papers

This is one of the most common tweets you'll see from scientists who aren't very active on Twitter. When a new paper is published, they'll log on, post the title of their paper with a link to the paper, and then log off until their next paper is published. There are so many more types of tweets that you can compose, but I'm going to walk through how you can lift your game with these types of tweets, which are one of the main reasons many scientists are on Twitter.

__1. Add an image from the paper to go with the tweet.__ Have a look through the paper to see if there's a striking image that can be used. If there are no figures in your paper, you can just take a screenshot of the abstract or a particularly interesting part of the paper. [Skitch](https://evernote.com/products/skitch) is a handy app for screenshots, as you can easily annotate and highlight images. If you're on your smartphone, you can also just take a screenshot and crop the image to your needs. You can add up to four images per tweet, so if you have more than one striking image you can take advantage of this

__2. Add a quote from the paper.__ Find a striking quote from the paper to include in your tweet. You can either write this as text, or take a screenshot of it and then highlight the quote, using Skitch or other software.

__3. Tag your co-authors and the journal.__ Co-authored papers are a team effort, so you should acknowledge your team. In addition, this gives your followers the chance to find more people to follow. If you have a lot of co-authors, remember that you can also tag authors in images as well. This means that your co-authors will get notified of the tweet and others can see who else is involved in the resesarch.

__4. Add your own commentary of the paper.__ If you're not including a quote, you should share *why* you think the paper is interesting. Maybe it's your first paper or it's part of a project you've been working on for a long time.

__5. Add a link to a non-paywalled version of the paper.__ People might see the tweet, but not bother clicking on the link if they do not have institutional access to the journal. However, if you include a link to a preprint or a postprint^[A preprint is a version of an article that is posted to a preprint repository before peer-review. Most journals allow you to submit papers that have been posted as preprints and do not consider this dual-publication. Check your journal's policies. A postprint is a version of paper that has undergone peer-review but has not been typeset. These are typically posted to the author's personal or institutional website. Postprints are usually just a PDF version of the final Word document that was sent to the journal. Unlike preprints, journal policies vary regarding postprints, so it's best to check before uploading one.], then people know they can easily access the paper.

In the following example of a paper I co-authored, I mention a summary of the results, highlight the sample size as this is a notable feature of the paper, tag my co-authors who are on Twitter, include a link to a postprint of the article, and attach a figure from the article.


```{r tweet-about-paper, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1161171300037079040")
```

Here's another good example, with [\@GuyProchilo](https://twitter.com/GuyProchilo) sharing his first first-author paper. Notice that Guy tags his co-authors, the journal, adds a hashtag (#IOpsych) relevant to his research field in this tweet (Industrial-organisational psychology), and includes a striking image of the paper.


```{r tweet-from-guy, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/GuyProchilo/status/1197095447757606919")
```

While single tweets announcing new papers can be an effective way to share your new work, I think a thread, which are a series of connected tweets, does a better job---see [Chapter 4](#advanced-twitter-skills) for how to make an effective thread to introduce your new work.

### Tweeting about other people's research papers

There are only so many articles that you can co-author. So in addition, you should also share papers that you find interesting. If you haven't already, you should set up new publication alerts using keywords that are relevant to your work. Pubmed and Google Scholar are popular options for setting up paper alerts. The advantage of Google Scholar is that it will also capture new preprints and theses. 

Below is an example of tweeting a link to a new paper, in which I added an image and quote from the paper, and mentioned the lead author and the journal that it was published in.

```{r tweet-about-paper-with-image, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1234808705666908162")
```


### Sharing information about yourself

Twitter provides a great way to connect with other researchers and to share information about yourself that you typically can't convey in a scientific paper. This is the sort of information that you could only normally learn in person, such as at a conference or a lab visit. But as travel is expensive and time-consuming, this option isn't open for everyone. 

Here's an example from [\@_DaniBeck](https://twitter.com/_DaniBeck). 

```{r tweet-introduction, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/_DaniBeck/status/1230448824277196800")
```

### Sharing your toolkit

As scientists, we use several tools in our day-to-day work. Even if these tools seem commonplace to you, your followers will appreciate learning about new ways to do their work. Here's an example of me tweeting about the ['bookdown' R package](https://bookdown.org/home/), which I've used to write this book. 

```{r tweet-toolkit, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1235114385569439744")
```

Notice that the link has been converted into a small image with information, which is called a [Twitter Card](https://developer.twitter.com/en/docs/tweets/optimize-with-cards/overview/abouts-cards). This is a feature of some websites (including this one), which include Twitter Card information in their underlying code. If you’re tweeting from your smartphone, the Twitter Card will appear (if it’s available) as soon as you enter the link. This doesn’t occur when tweeting from your desktop, but you can quickly check if there’s a Twitter Card associated with the link you’re going to post on the [Twitter Card Validator](https://cards-dev.twitter.com/validator) website. If the Twitter Card contains a nice image then you might not need to add your own. You can also delete the Twitter Card from your tweet as well.


### Asking a question

Twitter can be a great way to get advice. If you don't have many followers, you should also consider tagging experts that might know the answer (but don't spam people, so use this sparingly). Here's an example that I used, which was also used for researching this book.

```{r tweet-ask-question, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1234836938106003456")
```

This book is written for both HTML (online) and PDF (print) output. This is a book about Twitter, so I use a lot of tweet examples. One benefit of an online version is that you can embed tweets, which means that readers can visit the tweet's original source and read any replies to the tweet. It's also easy to see at a glance how many people have engaged with the tweet, as these numbers are attached to the bottom of embedded tweets. However, converting a HTML element into a image for a PDF is not as straightforward as you would think, as you essentially have to take a screenshot for each the tweet. To avoid this extra work, I put a call out on Twitter:

```{r tweet-call-out, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1275705042385940480")
```

Fortunately, Garrick Aden-Buie ([\@grrrck](https://twitter.com/grrrck)) found this tweet and updated his ['tweetmd' R package](https://github.com/gadenbuie/tweetrmd), which solved my problem. 

```{r tweet-tweetmd, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/grrrck/status/1276198597596459018")
```

### Answering questions

A great thing to tweet when you have have nothing to tweet is to help other people by answering their questions. There are several hashtags that are often used by scientists to ask questions, such as #AcademicChat and #PhDchat. What might seem like an easy solution to you may save someone else hours of work. 

### Replying to tweets

One of the fastest ways you can build connections on Twitter and to establish yourself as an expert in your area is to reply to other people's tweets. This seems counterintuitive, as one would think that conventional tweets are a better way to achieve these goals, but replies are almost guaranteed to be seen by the person you're replying to, whereas conventional tweets may not be seen at all (or quickly scrolled past), especially if you're new to Twitter. If others are following both yourself and the person you're replying to, they'll see the original tweet and your reply in their feed. 


### Sharing memes

Memes are a fun and effective way to demonstrate knowledge and authority in your research area. Notice in this example that this tweet was a quote retweet, so that people can see that I've remixed someone else's tweet.



```{r tweet-meme, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1234446547456266240")
```



### Sharing videos 

You can include a video that’s up to two minutes and twenty seconds long in your tweets. In the following tweet, I’ve included a link to an eleven-minute instructional YouTube video. I can’t show the entire video due to Twitter’s limits, so I’ve included a short preview to entice people to click on the link to the full video.


```{r tweet-video, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1162002047794864128")
```

If I was only to post a link to the video on YouTube, the Twitter Card would only show a static image. This is better than nothing, but you're much better off with a short preview. 

Also note that in this tweet I used the #Rstats hashtag, I tagged the author of the R package and tagged the software service I used.

People can’t always listen to audio when they’re scrolling Twitter, so it’s a good idea to include subtitles. This also increases accessibility for those with hearing impairments. When you upload your video to YouTube, it can [automatically transcribe your audio](https://support.google.com/youtube/answer/6373554). In my experience, the transcription is about 80% accurate, so you will need to do *some* editing of the transcription. However, this is much quicker than transcribing your video or audio clip from scratch. Another option for adding subtitles to your video is the ["Cliptomatic" app](https://www.apalon.com/clipomatic.html). Like YouTube, the transcription is about 80% accurate, but the editing function is pretty easy to use


## Document your experiences {.unnumbered}

I hardly ever sit down and think, "what will I tweet?". Most of my tweets are simply offcuts of work I'm already doing, or things I'm already discussing in real life with colleagues.

For example, I was recently having a chat with some colleagues about grant applications. I told them a principle I used to write my applications, which they found useful. So I turned this into a tweet, which also seemed to resonate on Twitter.



```{r tweet-grant, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1233838022224941061")
```

This thread is another example of ["showing my work"](https://austinkleon.com/show-your-work/). I didn't have to think of something new to tweet, as I essentially copy and pasted my reply email.



```{r tweet-research-design, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1150349430303989761")
```

Peer-review is a task that most of us do, so you can take this opportunity to make a wider commentary about your research topic. For instance, I recently shared a point about power analysis using a peer-review I was conducting.


```{r tweet-power-analysis, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1235109801543086082")
```

Let's break down this tweet:

1. I used a line break to clearly separate the two parts of this tweet.
2. The screenshot image is eye catching (especially the plot) and conveys a lot of information.
3. The 'down arrow' emoji adds a tiny bit of flair.
4. I tagged a tool that I mentioned [\@Jamovistats](https://twitter.com/Jamovistats), so that followers could easily find out more.
5. I numbered this tweet "1/2", to signal that this is a two-part tweet. My followers will see these two tweets back-to-back, but people who see this tweet as a retweet will now know there is more than one tweet on this topic, and can then check the full conversation.

Of course, if you're going to tweet about peer review you will need to keep things as anonymous as possible! Also consider that the author of the paper may follow you on Twitter or come accross your tweet if it gets retweeted.

You can also tweet about being at the *other end* of the peer review process.

```{r tweet-peer-review, echo=FALSE}
tweetrmd::include_tweet("https://twitter.com/dsquintana/status/1184586845935546370")
```

Remember, the key is sharing the process of your work, not just the final product.